---
title: "Data Import"
output: github_document
---

``` {r setup, echo=FALSE, message=FALSE}
library(tidyverse)
library(readxl) #read excel files
library(haven) # read sas
```



This document will show how to import data.


## Import the FAS Litters CSV

``` {r}
# Use this code to import the dataset.
litters_df = read_csv("data/FAS_litters.csv")
# Instead of importing the whole package, just take the one thing you need.
litters_df = janitor::clean_names(litters_df)
```

## Look at the dataset
``` {r}
litters_df

head(litters_df)
tail(litters_df, 10)
```

``` {r eval=FALSE}
# eval=FALSE Because popup confuses the R markdown
view(litters_df)
```

## Learning assessment

Load the FAS_pups dataset using the absolute path.
```{r, eval=FALSE}
pups_df = read_csv("Users/mayakrishnamoorthy/Documents/Columbia Classes/P8105/data-wrangling-i/data/FAS_pups.csv")
```

Load the FAS_pups dataset using the relative path.

``` {r}
pups_df = read_csv("data/FAS_pups.csv")
pups_df = janitor::clean_names(pups_df)
```

## Look at read_csv options

col_names and skipping rows
``` {r}
litters_df = 
  read_csv(
    file = "data/FAS_litters.csv"
    #col_names = FALSE,
    #skip=1
  )
```

What about missing data?

``` {r}
litters_df = 
  read_csv(
    file = "data/FAS_litters.csv",
    na = c("NA", "", ".")
  )

litters_df = janitor::clean_names(litters_df)
pull(litters_df, gd0_weight)
``` 


What if we code `group` as a factor variable?

``` {r}
litters_df = 
  read_csv(
    file = "data/FAS_litters.csv",
    na = c("NA", "", "."),
    col_types = cols(
      Group = col_factor()
    )
  )
```


## Import an excel file

Import MLB 2011 summary data.

``` {r}
mlb_df = read_excel("data/mlb11.xlsx", sheet = "mlb11")
```


## Import SAS data

``` {r}
pulse_df = read_sas("data/public_pulse_data.sas7bdat")
```

